---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

First project of reproducible reserch course week 2 by Rodrigo Tintor

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r load}
activity <- read.csv("activity.csv", na.strings = "NA")
activityOutNA <- activity[!is.na(activity["steps"]),]
activityOutNA["date"] = as.Date(activityOutNA$date)
summary(activityOutNA)
head(activityOutNA)
```

## What is mean total number of steps taken per day?

```{r mean}
total_act <- aggregate(x = activityOutNA$steps, FUN = sum, by = list(activityOutNA$date))
mean_act <- aggregate(x = activityOutNA$steps, FUN = mean, by = list(activityOutNA$date))
median_act <- aggregate(x = activityOutNA$steps, FUN = median, by = list(activityOutNA$date))
hist(total_act$x, , xlab = "Steps by date")
title(main = "Histogram of total steps per day")
```

## What is the average daily activity pattern?

```{r pattern}
total_act_time <- aggregate(x = activityOutNA$steps, FUN = sum, by = list(activityOutNA$interval))
plot(total_act_time$Group.1, total_act_time$x, type = "l", xlab = "Interval", ylab = "Steps", main = "Average step per interval per day")
```

## Imputing missing values
Missing values by column

```{r imputing}
naNumbers <- colMeans(is.na(activity)) * 100
naNumbers[naNumbers > 0]
```

For missing values we will use the median of the same interval

```{r imputing2}
median_act_inter <- aggregate(x = activityOutNA$steps, FUN = median, by = list(activityOutNA$interval))
activityNA <- activity[is.na(activity["steps"]),]
activityNA["date"] = as.Date(activityNA$date)
activityNA$steps <- median_act_inter$x[match(activityNA$interval, median_act_inter$Group.1)]

activityNoNA <- rbind(activityOutNA, activityNA)

total_act_f <- aggregate(x = activityNoNA$steps, FUN = sum, by = list(activityNoNA$date))
mean_act_f <- aggregate(x = activityNoNA$steps, FUN = mean, by = list(activityNoNA$date))
median_act_f <- aggregate(x = activityNoNA$steps, FUN = median, by = list(activityNoNA$date))
hist(total_act_f$x, , xlab = "Steps by date", main = "Average step per interval per day (filled NA)")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays}
w1 <- weekdays(activityNoNA$date, abbreviate = TRUE)
w2 <- w1=="Sat" | w1 =="Sun"
week <- factor(w2, labels = c("weekday","weekend"))
activityNoNA["daytype"] <- week
activityNoNAWD <- subset(activityNoNA ,daytype=="weekday")
activityNoNAWE <- subset(activityNoNA ,daytype=="weekend")
mean_act_f2 <- aggregate(x = activityNoNAWD$steps, FUN = mean, by = list(activityNoNAWD$interval))
plot(mean_act_f2$Group.1, mean_act_f2$x, type = "l", xlab = "Interval", ylab = "Steps", main = "Average step per interval per weekday")
mean_act_f3 <- aggregate(x = activityNoNAWE$steps, FUN = mean, by = list(activityNoNAWE$interval))
plot(mean_act_f3$Group.1, mean_act_f3$x, type = "l", xlab = "Interval", ylab = "Steps", main = "Average step per interval per weekend")
```

Thanks in advance for your comments
Rodrigo

